On the properties of discs around accreting brown dwarfs

### 2010-05-27 BD_disc/paper.tex
Best copy up to date for TJH to check and submit
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@2288 25464f21-6099-402c-b30a-640f1d40d38b

As shown in Figures \ref{flare_-12_183} and \ref{flare_-7_193} and
discussed in Section \ref{disc_flaring} BDD systems under vertical
hydrostatic equilibrium have highly flared discs. As discussed in
\cite{walker_2004} this increased flaring (when compared to CTTS
stars) leads to occultation at the star at lower inclination angles
and, therefore, significant changes to the SED. Increases in the
inclination angle for these systems quickly lead to a significant
proportion of the stellar flux being intercepted and reprocessed by
the highly flared disc. This reprocessing will lead to a change in the
flux levels at the shorter, bluer, wavelengths as more stellar flux is
intercepted by the disc. It will also lead to significant changes in
the flux reaching the observer from the inner and outer regions of the
disc as the system approaches edge on. Also, as discussed in Section
\ref{disc_struct} the addition of dust sublimation leads to a change
in the shape of the inner edge, where the temperature is
sufficient. Figures \ref{no_sub_183_a} to \ref{sub_179_a}
\ref{sub_181_a} show a range in inner edge shapes, from flat walls
through concave to convex curves, caused by the radial density profile
of the disc and the dependence of the sublimation temperature on
density. This change in shape, as noted for Herbig Ae stars by
\cite{tannirkulam_2007} XXXCHECK FOR A CTTS REFERENCESXXX, will lead
to changes in the characteristics of the SED, or derived IR excess,
with inclination angle \citep{tannirkulam_2007}, for close to face on
viewing angles.

@@

The changes in flux with inclination will clearly lead to changes in
the observed magnitudes and colours with inclination. Figure
\ref{flare_inc_mag} shows the $M_V$ and $M_J$ magnitudes in the top
and bottom panels respectively. The magnitudes are then marked as
crosses as a function of inclination. The systems shown in both panels
have an age of 1 Myr, mass of $0.04M_{\odot}$, rotation rates of 0.5
and an areal coverage of 10\%. The systems with accretion rates of
\logmdot = $-$12, $-$9 and $-$7 are shown in black, red and blue,
respectively. The panels also show the system with the highest
accretion rate but with a rotation rate of 5 days as a dashed
line. The vertical dotted lines in the lowest panel simply illustrate
the total change in magnitude, over the shown inclination range.

\begin{figure*}
  \centering
  \includegraphics[scale=0.6,angle=90]{./Fig/flare_inc_mag.ps}
  \caption{Figure showing $M_V$ and $M_J$ magnitudes as a function of
    inclination, shown as top, middle and lower panels
    respectively. The magnitudes of systems with $M_*=0.04M_{\odot}$,
    an age of 1 Myr, areal coverage of 10\% and rotation rate of 0.5
    days are shown as crosses. The systems with accretion rates of
    \logmdot = $-$12, $-$9 and $-$7, are shown in black, red and blue
    respectively. The panels also the systems with the highest
    accretion rate but for a rotation rate of 5 days. The lowest panel
    also shows, as horizontal dotted lines, the total change of colour
    with the illustrated inclinations. \label{flare_inc_mag}}
\end{figure*}

Accretion flux was shown to dominate the underlying, intrinsic,
photospheric SED for accretion rates of \logmdot = $-$9 (see Figure
\ref{acc_dom}). As one would expect this leads to significant changes
in the derived photometric magnitudes for bands blueward of a few
microns, where the accretion and photospheric flux dominate. This is
shown in the top panel of Figure \ref{flare_inc_mag}. As the accretion
rate increases the system moves to brighter magnitudes in $M_V$. The
move to brighter magnitudes becomes significant once the accretion
rate exceeds \logmdot = $-$9. Again moving from the fast to slow
rotation period increases the potential energy released and in this
case results in a brighter $M_V$ magnitude. As one moves to longer
wavelength photometric bands, as shown in the middle panel. The
increase in accretion rate affects these magnitudes, i.e. $M_J$ are
less significantly affected. Figure \ref{flare_inc_mag} also shows the
change in magnitude caused by obscuration from the flared outer
disc. The occultation starts earlier for the higher accretion rates,
with the magnitudes in $M_V$ and $M_J$ moving fainter at earlier
inclinations for the higher accretion rates.

XXXTHIS SECTION IS ON THE INNER EDGE AND NEEDS SORTINGXXX
179 
221

Figure \ref{inner_inc} shows the SEDs for the viewing angles of 0, 27,
39, and 48 $^{\circ}$, close to face on. The top and bottom panels of
Figure \ref{inner_inc} show SEDs for models with a curved, as
exemplified by Figure \ref{flare_-7_193}, and flat inner wall, as
shown in Figure \ref{flare_-12_183} respectively. The specific models
chosen have been selected to minimise flux differences between the two
SEDs but have distinct inner edges. XXXMODELSXXX Comparing the top and
bottom panel show that flux levels for the higher accretion rate, with
the larger and curved inner edge, vary less distinctly with
inclination angle than the systems with a vertical wall.

@@

The $(J-K)_0$ colour of the higher accretion rate system changes less
as a function of inclination when compared to the lower accretion rate
system. A total change in $(J-K)_0$ of $\approx$ XXXVALUESXXX are
found for the model with the flat and curved inner edge
respectively. This reduction in the change in IR colour as a function
of inclination for viewing angles close to face-on, is caused by the
larger (vertically) and convex inner wall in the extreme accreting
case. For the convex inner wall the effective emitting area one
observes does not change as much as a vertical wall as a function of
inclination, leading to this reduced dependence of IR colour on
inclination.

@@

Figure \ref{disc_frac_mips} shows that as we increase the wavelength
even further, into the range of the MIPS photometric bands, the
seperation between naked and BDD systems increases still further. The
inset panels show that the seperation between the naked and BDD
systems is larger than in Figure \ref{disc_frac_irac}. Additionally,
there are clear correlations of MIPS positions with accretion rate and
inclination. For the extreme accretors the systems are clearly
delineated by accretion rate and inclination. This is as the emission
is coming from greater radial positions within the disc where the
structure of the disc is a finer function of the input
variables. XXXNJM TO CHECK THIS MORE NEEDEDXXX

>>>>

As shown in Figures \ref{flare_-12_183} and \ref{flare_-7_193} and
discussed in Section \ref{disc_flaring} BDD systems under vertical
hydrostatic equilibrium have highly flared discs. As discussed in
\cite{walker_2004} this increased flaring (when compared to CTTS
stars) leads to occultation at the star at lower inclination angles
and, therefore, significant changes to the SED. Increases in the
inclination angle for these systems quickly lead to a significant
proportion of the stellar flux being intercepted and reprocessed by
the highly flared disc. This reprocessing will lead to a change in the
flux levels at the shorter, bluer, wavelengths as more stellar flux is
intercepted by the disc. It will also lead to significant changes in
the flux reaching the observer from the inner and outer regions of the
disc as the system approaches edge on. Also, as discussed in Section
\ref{disc_struct} the addition of dust sublimation leads to a change
in the shape of the inner edge, where the temperature is
sufficient. Figures \ref{no_sub_183_a} to \ref{sub_179_a}
\ref{sub_181_a} show a range in inner edge shapes, from flat walls
through concave to convex curves, caused by the radial density profile
of the disc and the dependence of the sublimation temperature on
density. This change in shape, as noted for Herbig Ae stars by
\cite{tannirkulam_2007}, will lead to changes in the characteristics
of the SED, or derived IR excess, with inclination angle
\citep{tannirkulam_2007}, for close to face on viewing angles.

@@

The changes in flux with inclination will clearly lead to changes in
the observed magnitudes and colours with inclination. Figure
\ref{flare_mag} shows the $M_V$ and $M_J$ magnitudes in the top
and bottom panels respectively. The magnitudes are then marked as
crosses as a function of inclination. The systems shown in both panels
have an age of 1 Myr, mass of $0.04M_{\odot}$, rotation rates of 0.5
and an areal coverage of 10\%. The systems with accretion rates of
\logmdot = $-$12, $-$9 and $-$7 are shown in black, red and blue,
respectively. The panels also show the system with the highest
accretion rate but with a rotation rate of 5 days as a dashed
line. The vertical dotted lines in the lowest panel simply illustrate
the total change in magnitude, over the shown inclination range.

\begin{figure*}
  \centering
  \includegraphics[scale=0.6,angle=90]{./Fig/flare_mag.ps}
  \caption{Figure showing $M_V$ and $M_J$ magnitudes as a function of
    inclination, shown as top, middle and lower panels
    respectively. The magnitudes of systems with $M_*=0.04M_{\odot}$,
    an age of 1 Myr, areal coverage of 10\% and rotation rate of 0.5
    days are shown as crosses. The systems with accretion rates of
    \logmdot = $-$12, $-$9 and $-$7, are shown in black, red and blue
    respectively. The panels also the systems with the highest
    accretion rate but for a rotation rate of 5 days. The lowest panel
    also shows, as horizontal dotted lines, the total change of colour
    with the illustrated inclinations. \label{flare_mag}}
\end{figure*}

Accretion flux was shown to dominate the underlying, intrinsic,
photospheric SED for accretion rates of \logmdot = $-$9 (see Figure
\ref{acc_dom}). As one would expect this leads to significant changes
in the derived photometric magnitudes for bands blueward of a few
microns, where the accretion and photospheric flux dominate. This is
shown in the top panel of Figure \ref{flare_mag}. As the accretion
rate increases the system moves to brighter magnitudes in $M_V$. The
move to brighter magnitudes becomes significant once the accretion
rate exceeds \logmdot = $-$9. Again moving from the fast to slow
rotation period increases the potential energy released and in this
case results in a brighter $M_V$ magnitude. As one moves to longer
wavelength photometric bands, as shown in the middle panel. The
increase in accretion rate affects these magnitudes, i.e. $M_J$ are
less significantly affected. Figure \ref{flare_mag} also shows the
change in magnitude caused by obscuration from the flared outer
disc. The occultation starts earlier for the higher accretion rates,
with the magnitudes in $M_V$ and $M_J$ moving fainter at earlier
inclinations for the higher accretion rates.

XXXTHIS SECTION IS ON THE INNER EDGE AND NEEDS SORTINGXXX

Figure \ref{inner_inc} shows the SEDs for the viewing angles of 0, 27,
39, and 48 $^{\circ}$, close to face on. The top and bottom panels of
Figure \ref{inner_inc} show SEDs for models with a curved, as
exemplified by Figure \ref{flare_-7_193}, and a flat inner wall, as
shown in Figure \ref{flare_-12_183} respectively. The specific models
chosen have been selected to minimise flux differences between the two
SEDs but have distinct inner edges. The top panel has a mass of
0.04$M_{\odot}$, \logmdot = -8, areal coverage of 1\%, a rotational
period of 5 days and an age of 10 Myrs. The bottom panel has a the
same model but with rotation period of 0.5 days and an age of 1
Myrs. Comparing the top and bottom panel show that flux levels for the
higher accretion rate, with the larger and curved inner edge, vary
less distinctly with inclination angle than the systems with a
vertical wall. Although this effect is marginal. The weakness of the
effect is probably due to mismatches in the disc structure and flux
levels. Ideally, we would take the curved inner edge model and replace
the inner edge with a vertical wall.

@@

The $(J-K)_0$ colour of the higher accretion rate system changes less
as a function of inclination when compared to the lower accretion rate
system. A total change in $(J-K)_0$ of $\approx$ 0.6 and 0.05 are
found for the model with the flat and curved inner edge
respectively. This reduction in the change in IR colour as a function
of inclination for viewing angles close to face-on, is caused by the
larger (vertically) and convex inner wall in the extreme accreting
case. For the convex inner wall the effective emitting area one
observes does not change as much as a vertical wall as a function of
inclination, leading to this reduced dependence of IR colour on
inclination.

@@

Figure \ref{disc_frac_mips} shows that as we increase the wavelength
even further, into the range of the MIPS photometric bands, the
seperation between naked and BDD systems increases still further. The
inset panels show that the seperation between the naked and BDD
systems is larger than in Figure \ref{disc_frac_irac}. Additionally,
there are clear correlations of MIPS positions with accretion rate and
inclination. For the extreme accretors the systems are clearly
delineated by accretion rate and inclination. This is as the emission
is coming from greater radial positions within the disc where the
structure of the disc is a finer function of the input
variables.
